Verifying Demonstrations

### 2020-07-03 paper/DB-Schnorr-nbit.tex
Cuts bit-by-bit version

\subsection{Security of the \(l\) bits}

Before we proceed to other security properties, we will address the \(l\)-bit 
problem:
what advantage does the adversary get from those \(l\) leaked bits of the 
second response?

\paragraph*{The unconditional \(l\)-bit problem}

We run the protocol \(n\) times and let the adversary learn \(m+l\) bits of the 
result~\(s_i, s_i'\).
This yields the following equation system:
\begin{align*}
  s_0[0]2^0 + \dotsb + s_0[m-1]2^{m-1} &= r_0 + c\alpha &\pmod q\\
  s_0'[0]2^0 + \dotsb + s_0'[m-1]2^{m-1} &= r_0 + c'\alpha &\pmod q\\
  \vdots \\
  s_{n-1}[0]2^0 + \dotsb + s_{n-1}[m-1]2^{m-1} &= r_{n-1} + c\alpha &\pmod q\\
  s_{n-1}'[0]2^0 + \dotsb + s_{n-1}'[m-1]2^{m-1} &= r_{n-1} + c'\alpha, &\pmod 
  q
\end{align*}
where \(m+l\) of \(s_i[j]\)'s and \(s_i'[j']\)'s for each \(i\) are known by 
the adversary.

\begin{lemma}\label{l-bit-advantage}
  If \[l < m-1+\frac{1}{n}-\frac{f(\lambda)}{n},\] then there are at least 
  \(2^{f(\lambda)}-1\) unique solutions, for some function~\(f\) of the 
  security parameter~\(\lambda\).
\end{lemma}

The proof is in \cref{proofs-lbit}.

\paragraph*{The \(l\)-bit problem with commitments}

The adversary has a commitment to the randomness~\(r_i\) in the form of \(R_i = 
g^{r_i}\) and a commitment to \(\alpha\) in the form of \(A = g^\alpha\).
This means that
\begin{align*}
  R = A^{-c'} g^{s'} &= A^{-c'} g^{s'[0] 2^0 + \dotsb + s'[m-1] 2^{m-1}} \\
                     &= A^{-c'} g^{s'[0] 2^0} \dotsb g^{s'[m-1] 2^{m-1}} \\
                     &= A^{-c'} g^{s'_{\text{known}}} g^{s'_{\text{unknown}}}.
\end{align*}
We will now show how the hardness of finding \(s'_{\text{unknown}}\) relates to 
the hardness of finding discrete logarithms.

@@

The proof is in \cref{proofs-lbit}.

As mentioned above, both \(l\) and \(k\) are logarithmic in the security 
parameter.

\subsection{\Acl*{ZK} and \acl*{PK}}\label{nbit-ZKPK}

@@

This is \emph{computational} \acl{ZK}:
The simulator chooses \[
  s_0, s_1\rgets \ZZ_q; \quad
  c_0, c_1\rgets \bin^k; \quad
  b\rgets \bin.
\]
Then it can set \[
  R_0\gets g^{s_0} A^{c_0}, \quad
  R_1\gets g^{s_1} A^{c_1}.
\]
Now the simulator must come up with those \(l\) bits.
\Cref{l-bit-advantage,DBSPtoDLP} say that distinguishing these bits from random 
is as hard as the \ac{DLP}.
So, the simulator can simply choose the \(l\) bits~\(L\rgets \bin^l\) uniformly 
at random.

@@

%Therefore the standard proof for Schnorr as a malicious-verifier \ac{ZKPK} 
%protocol with soundness error \(2^{-k n}\) still 
%holds~\cite[\eg][]{OnSigmaProtocols}.
%(We note that \(l\), which is also the length of the challenge bit string, must 
%be logarithmic in the security parameter, \(\lambda\), for a malicious 
%verifier.)

\subsection{\acs*{DBMF}, \acs*{DBTF} and \acs*{DBDH} resistance}

\paragraph{\Acl{DBMF}}

In \ac{DBMF}, the adversary will try to authenticate to a verifier by querying 
a far-away prover.
By \cref{TaintedMF}, pure relaying will be detected by the verifier.
So the adversary must resort to things like querying in advance \etc.

\begin{theorem}[\acs*{DBMF} resistance]
  Let \(\adv\) be a \((t, \qobs, \qp, \qv)\)-MF adversary, then \(
    \Adv_{\DBS}^{\text{MF}}(\adv) =
      \left(\frac{1}{2} + \frac{1}{2^{|q|-l}}\right)^n,
  \) where \(m, q\) are as in \cref{DB-Schnorr-nbit}.
\end{theorem}

\begin{proof}
  It follows from the \ac{ZK} property (\cref{nbit-ZKPK}) that the adversary 
  cannot learn anything by observation.
  This leaves the adversary with the following:
  \begin{enumerate}
    \item guess which challenge the verifier will use and request the correct 
      response from the prover, receiving the challenge first would taint the 
      session (\cref{TaintedMF});
    \item guess the second response.
  \end{enumerate}

  The first case yields success with probability \(\frac{1}{2}\), since the 
  choice is uniformly distributed over the two challenges.
  The adversary requests all \(|q|\) bits of the response~\(s_{\hat b}\) and 
  \(l\) bits of \(s_{1-\hat b}\).
  When the verifier sends challenge bit \(b_i = \hat b\), the adversary replies 
  with a bit from \(s_{\hat b}\).
  If \(b_i = 1-\hat b\), then the adversary uses the \(l\) bits.

  If the adversary guessed \(\hat b = b\) correctly, then the verifier will 
  accept.
  Otherwise, the adversary must compute the remaining \(|q|-l\) bits of 
  \(s_{1-\hat b}\).

  Guessing the remaining bits yield a success probability of 
  \(\frac{1}{2^{|q|-l}}\).
  \Cref{DBSPtoDLP} states that any algorithm that can compute the remaining 
  bits can also solve a corresponding \ac{DLP}.
  It follows that the adversary cannot do non-negligibly better than guessing 
  the remaining bits.

  Thus, the success probability is \(\frac{1}{2}+\frac{1}{2^{|q|-l}}\) per 
  round, or \(\left(\frac{1}{2}+\frac{1}{2^{|q|-l}}\right)^n\) in total.
\end{proof}

\paragraph{\Acl{DBTF}}

We cannot achieve \ac{DBTF} resistance.
Indeed, the prover can send \(l+\nu\) bits of both response strings, \(s_0\) 
and \(s_1\), to the adversary.
When the adversary receives the \(l+1\)th bit, he learns \(b\).
He can send \(b\) to the prover and answer correctly to the verifier, provided 
\(\nu \geq 1\).
By the time he receives the next challenge from the verifier, the prover will 
have sent him the remaining bits of \(s_b\) without revealing more bits about 
\(s_{1-b}\).
This would not taint the session.
As long as \(l+\nu\) is sufficiently small, the adversary cannot compute the 
remaining bits of \(s_{1-b}\) to compute the prover's secret.

\paragraph{\Acl{DBDH}}

The protocol is also secure against distance hijacking due to the fact that the 
authenticating bit string is used during the \ac{DB} phase.
(Brands-Chaum used the challenge bit string, something that made their protocol 
vulnerable.)
\Ac{DBDH} requires that the adversary finds a collision between his response 
and that of the honest prover.
Thus the probability of success is equivalent to a collision for the responses 
for the chosen challenge --- in each round.

\begin{theorem}[\acs*{DBDH}/\acs*{DBDF} resistance]
  Let \(\adv\) be a \((t, \qobs, \qp, \qv)\)-DH adversary, then 
  \(\Adv_{DBS}^{\text{DH}}(\adv) = \frac{1}{2^n}\).
\end{theorem}

\begin{proof}
  The honest prover will choose \(\rho\) such that \(R = g^\rho\).
  The adversary can send his own \(R' = g^{\rho'}\) to the verifier, or let 
  \(R' = R\) to not interfere.

  The verifier will choose \(c_0, \dotsc, c_1\).
  The honest prover will compute \(s_i = \rho - c_i \alpha\), where \(\alpha\) 
  is his secret.
  If the adversary chose \(R' = g^{\rho'}\), he can compute \(s_i' = \rho' - 
  c_i \alpha'\), where \(\alpha'\) is the adversary's secret.
  If the adversary didn't choose \(R'\), he can only guess the values for 
  \(s_i'\).

  During the fast phase, the adversary has two options:
  for each challenge bit, the adversary can \(\Commit\) to a response bit or to 
  \(\Prompt\) the honest prover to send a bit.

  If the adversary chose \(R'\), then he knows \(\rho'\) and can compute the 
  correct \(s_i' = \rho' - c_i \alpha'\) for \(i\in \bin\).
  If the adversary \(\Commit\)s:
  Now, the adversary must commit to one bit of \(s_0'\) or \(s_1'\) before 
  seeing the verifier's choice \(b_i\).
  He can make a guess~\(\hat b\) at the verifier's choice~\(b\).
  Then he commits to the next bit of \(s_{\hat b}'\).
  If the challenge~\(b_i \neq \hat b\), he will commit to the same bit for the 
  next round.
  If indeed \(\hat b = b\), the verifier will accept.
  The success probability is \(\frac{1}{2}\) over all bits.

  If the adversary \(\Prompt\)s:
  The prover computed \(s_0\) and \(s_1\) using \(\rho\) and \(\alpha\).
  However, the verifier will use \(R'\) and \(A'\) in the verification step.
  This will succeed if \(s_b = s_b'\), the probability of which is 
  \(\frac{1}{q}\).
  This means that the probability of success for prompting a bit is 
  \(\frac{1}{2}\), or \(\frac{1}{2^{|q|}}\) over all bits.

  If the adversary didn't choose \(R'\) he consequently doesn't know \(\rho\) 
  and cannot compute the correct \(s_0\) and \(s_1\).
  Thus, committing yields \(\frac{1}{2}\) probability per bit.
  The same applies for prompting.

  Obviously, the better strategy for the adversary is to choose \(R' = 
  g^{\rho'}\).
  It's also a better strategy to \(\Commit\) (\ac{DBDF}) rather than 
  \(\Prompt\) (\ac{DBDH}).
  We thus have that
  \(\Adv_{\DBS}^{\text{DH}}(\adv) = \frac{1}{2^n}\) for \(n\) rounds,
  which concludes the proof.
\end{proof}

>>>>

\subsection{Security analysis}

We can prove this protocol secure against \ac{DBMF}, \ac{DBDH} and \ac{DBDF}; 
but not against \ac{DBTF}.
Those proofs are available in the full version of the paper.
Here, we will discuss only the most crucial differences.

\paragraph*{The \(l\)-bit problem}

We can show that the system of linear equations created by the protocol and 
consisting of leaked bits from the second reply is underdetermined.
However, we cannot get information theoretic security, that would require 
leaking no bits, but we leak some bits.
The only lead an adversary get to those remaining unknown bits are the 
commitments.
The adversary has a commitment to the randomness~\(r_i\) in the form of \(R_i = 
g^{r_i}\) and a commitment to \(\alpha\) in the form of \(A = g^\alpha\).
This means that
\begin{align*}
  R = A^{-c'} g^{s'} &= A^{-c'} g^{s'[0] 2^0 + \dotsb + s'[m-1] 2^{m-1}} \\
                     &= A^{-c'} g^{s'[0] 2^0} \dotsb g^{s'[m-1] 2^{m-1}} \\
                     &= A^{-c'} g^{s'_{\text{known}}} g^{s'_{\text{unknown}}}.
\end{align*}
We will now show how the hardness of finding \(s'_{\text{unknown}}\) relates to 
the hardness of finding discrete logarithms.

@@

The proof of this lemma can be found in the full version of the paper.

\paragraph*{\Acl*{ZK} and \acl*{PK}}\label{nbit-ZKPK}

@@

The change is with the \acl{ZK} property.
This is \emph{computational} \acl{ZK}:
The simulator chooses \[
  s_0, s_1\rgets \ZZ_q; \quad
  c_0, c_1\rgets \bin^k; \quad
  b\rgets \bin.
\]
Then it can set \[
  R_0\gets g^{s_0} A^{c_0}, \quad
  R_1\gets g^{s_1} A^{c_1}.
\]
Now the simulator must come up with those \(l\) bits.
By \cref{DBSPtoDLP} and some lemmas in the full version of the paper, we can 
see that distinguishing these bits from random is comparable to solving the 
\ac{DLP}.
So, the simulator can simply choose the \(l\) bits~\(L\rgets \bin^l\) uniformly 
at random.

@@

\paragraph*{Why not \acs*{DBTF}?}

The problem is that asymptotically, \(l\) and \(l+1\) are rather close.
So the colluding parties can exchange \(l+1\) bits instead of just \(l\).
This way they can create a buffer of responses.
However, this means that they can stretch the distance bound; at some point, 
\(l+\epsilon\) will be too large and make \(\alpha\) easy enough to guess.
So the incentive changes from the legitimate party not wanting to cheat, into 
the legitimate party not wanting to cheat \enquote{too much}.
However, we have not formally treated this part more than concluding that we 
cannot prove \ac{DBTF} resistance in terms of \cref{TaintedTF,TFresistance}.
